---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 10,
  fig.height = 8
)
```

# cyclic.coefs

<!-- badges: start -->
<!-- badges: end -->

**cyclic.coefs** provides tools for exploring series with periodic coefficients through dot products. The package implements the mathematical pattern:

$$\sum_{i=1}^{n} f_i \cdot g_i$$

where **f** is a periodic sequence (sine, cosine, triangle wave, etc.) and **g** is a base sequence (harmonic, geometric, exponential, etc.). This framework is fundamental to Fourier analysis and series expansions.

## Installation

Install the development version from GitHub:
``` r
# install.packages("pak")
pak::pak("randomplayables/cyclic.coefs")
```

## Quick Start

Combine a sine wave with exponential decay:

```{r example}
library(cyclic.coefs)

# Create a combination
res <- combine(
  .f = \(x) sin_cyc(),
  .g = \(x) exp_decay_g(rate = 0.5)
)

# View results
print(res)
```


Visualize the interaction:

```{r plot, fig.height=6}
plot(res)
```


## Key Functions

**Periodic coefficients (.f functions):**
- `sin_cyc()`, `cos_cyc()` - Trigonometric sequences
- `triangle_cyc()`, `sawtooth_cyc()`, `square_cyc()` - Geometric waves
- `pulse_cyc()` - Rectangular pulse waves
- `complex_cyc()` - Complex exponentials for Fourier analysis

**Base sequences (.g functions):**
- `harmonic_g()` - Harmonic series (1/n)
- `power_g()` - Power sequences (n^p)
- `geometric_g()` - Geometric sequences (r^n)
- `exp_decay_g()` - Exponential decay
- `fibonacci_g()`, `binomial_g()`, and more

## Classic Example: Fourier-like Series

```{r fourier}
# Sine with harmonic coefficients: Σ(1/n)sin(nx)
fourier_like <- combine(
  .f = \(x) sin_cyc(seq(0, 2*pi, length.out = 33)[-33]),
  .g = \(x) harmonic_g(1:32)
)

summary(fourier_like)
```
